---
title: 'The Impact of ESG Components on equity performance : A Comparative Analysis
  across Indexes and Sectors '
author: "William"
date: "2023-08-12"
output:
  html_document:
    df_print: paged
  word_document:
    fig_caption: yes
    reference_docx: Report.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(readxl)
library(visdat)
library(knitr)
#library(kableExtra)
library(summarytools)
library(pastecs)
library(corrplot)
library(flextable)
library(gt)
```

# Data Cleaning

```{r}
#import data
data=as.data.frame(read_excel('mergeddata  (1).xlsx'))#merged data) 
```

```{r}
# Data cleaning
# Structure of the data
glimpse(data)

#checking for duplicated data
anyDuplicated(data)

# Identify stock tickers for each company in 2020
stock_2010 = data %>%
  filter(Year == 2020) %>%
  select(Stock, Name)

# Filter out the duplicates in 2021 based on 2010 data
data = data %>%
  left_join(stock_2010, by = "Name", suffix = c("", "_2020")) %>%
  mutate(keep_row = if_else(Year != 2021 | (Year == 2021 & Stock == Stock_2020), TRUE, FALSE)) %>%
  filter(keep_row) %>%
  select(-Stock_2020, -keep_row)


# checking fr missing values
anyNA(data)
#Visualizing missing data
vis_miss(data, facet = Year)

data %>%
  group_by(Year) %>%
  summarise(across(everything(), ~sum(is.na(.x)), .names = "missing_in_{.col}"))
# missing values per column
colSums(is.na(data))
# For reliability we will eliminate all missing values in the data
data_clean = data %>%
  filter(!is.na(`Close price`) & !is.na(`Refinitiv - ESG Score`))

# Mapping stock indices to countries
data_clean$Country = ifelse(data_clean$Stock == 'SPX', 'USA',
                            ifelse(data_clean$Stock == 'FTSE', 'UK',
                                   ifelse(data_clean$Stock == 'N225', 'Japan', NA)))

char = c("Stock", "Year", "TRBC Business Sector", "Country")
data_clean[char] = lapply(data_clean[char], factor)

```



# Descriptive Exploration

```{r}
a=data_clean %>% 
  select(Stock, Year) %>% 
  table#%>% 
  
kable(a)
```


```{r}
data_clean %>% 
  select(`TRBC Business Sector`, Year) %>% 
  table %>% kable()
  
```



```{r}
data_clean %>% 
  select(where(is.numeric)) %>% 
  descr(stats = c("mean", "sd", "min", "max","skewness"), transpose = TRUE) %>% 
  round(4) %>% 
  data.frame(`Variable Name`=row.names(.)) %>%
  select(c(6,1:5)) %>% 
  regulartable() %>% 
  bg(bg = "#000080", part = "header") %>%
  color(color = "white", part = "header")

```

```{r}
data_clean %>%
  ggplot(aes(`Close price`, fill = Stock)) +
  geom_density(alpha = 0.5)+
  labs(x = "Cosing Price", title = "Density Plot of Stock Close Prices") +
  theme_bw() +
  scale_fill_manual(values = c("green", "blue", "red"))
```

```{r}
# controlling outliers using log
data_clean %>%
  ggplot(aes(log(`Close price`), fill = Stock)) +
  geom_density(alpha = 0.5)+
  labs(x = "Log Cosing Price", title = "Density Plot of log Close Prices") +
  theme_bw() +
  scale_fill_manual(values = c("green", "blue", "red"))
```


```{r}
data_clean %>%
  group_by(Stock, Year) %>%
  summarize(Sum_Close_Price = sum(`Close price`)) %>% 
  ggplot(aes(x = Year, y = Sum_Close_Price, group = Stock, color = Stock)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", y = "Sum of Close Price", 
       title = "Close Prices by Year") +
  scale_color_manual(values = c("green", "blue", "red"))+
  theme_bw()

data_clean %>%
  group_by(Stock,Year) %>%
  summarize(Mean_Close_Price = mean(`Close price`, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = Mean_Close_Price, fill = Stock)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Stock", y = "Mean Close Price", 
       title = "Mean Close Prices by Stock") +
  theme_bw() +
  scale_fill_manual(values = c("green", "blue", "red"))


data_clean %>%
  ggplot(aes(Year, `Close price`, fill = Stock)) +
  geom_boxplot()+
  labs(x = "Year", title = "Close Prices by Stock") +
  theme_bw() +
  scale_fill_manual(values = c("green", "blue", "red"))

```
```{r}
data_clean %>% 
  ggplot(aes(`Refinitiv - ESG Score`, fill = Stock, 
             fill = Stock)) +
  geom_density(alpha = 0.2) +
  labs(title = "Density plot of ESG Score") +
  theme_bw()
```


```{r}
data_clean %>%
  group_by(Stock, `TRBC Business Sector`) %>%
  summarize(mean_Close_Price = mean(`Close price`)) %>% 
  ggplot(aes(x =`TRBC Business Sector`,y = mean_Close_Price,fill = Stock)) +
  geom_bar(stat = 'identity') +
  labs(x = "Year", y = "Average Close Price", 
       title = "Close Prices by Industry") +
  scale_fill_manual(values = c("green", "blue", "red"))+
  coord_flip()+
  theme_bw()
```


```{r}

data_clean %>% 
  ggplot(aes(x = `Refinitiv - ESG Score`, y = `Close price`, 
             color = Stock)) +
  geom_point() +
  labs(x = "ESG Score", y = "Close Price", 
       title = "ESG Combined Score vs. Close Price") +
  scale_color_manual(values = c("green", "blue", "red"))+
  theme_bw()

```

As shown from the plot the higher return was generated by FTSE and SPX companies. 

```{r}
data_clean %>%
  group_by(Stock, Year) %>%
  summarize(Mean_5Year_Return = mean(`Refinitiv - ESG Score`)) %>%
  ggplot(aes(x = Year, y = Mean_5Year_Return, fill = Stock)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Stock", y = "ESG Score", title = "ESG Combined Score by Stock") +
  theme_bw()+
   scale_fill_manual(values = c("green", "blue", "red"))

```


```{r}
data_clean %>% 
  select(ends_with('score')) %>% 
  vis_cor(cor_method = "pearson")

data_clean %>% 
  select(c(5,20:25))%>% #cor(use = "complete.obs")
  vis_cor(cor_method = "pearson")
```

There is a significant relationship between ESG scores and stock performance in the US, UK, and Japan
```{r}
cor.test(data_clean$`Refinitiv - ESG Score`, data_clean$`Close price`,
         method = 'spearman')
```
How do ESG scores impact stock performance in the US, UK, and Japan, and does this impact vary across different sectors

```{r}
aov(`Refinitiv - ESG Score` ~ Country + Year, data = data_clean) %>% 
  summary()
```
The relationship between ESG scores and stock performance varies across different sectors within each country

```{r}
aov(cbind(`Refinitiv - ESG Score`,`Close price`) ~ Country , data = data_clean) %>% 
  summary()
```
Do companies with higher ESG scores demonstrate better financial resilience during economic downturns & upturns compared to those with lower ESG scores within each industry in the S&P 500, FTSE 250, and N225?

```{r}
# Higher standardized ESG score >=mean
# Higher standardized ESG score <mean
avrg=mean(data_clean$`Refinitiv - ESG Score`)

data_clean$ESG_Level = ifelse(data_clean$`Refinitiv - ESG Score` >= avrg,
                              'High','Low') %>%
  as.factor()

data_clean$Upturn = ifelse(as.numeric(data_clean$Year)>2,
                           'Upturn','Downturn') %>% as.factor()
```


To address this research question, the subsequent hypotheses were crafted:
H2: Companies with higher ESG scores exhibit greater financial resilience during economic downturns compared to those with lower scores within the S&P 500, FTSE 250, and N225.

```{r}
data_clean %>%
  group_by(Stock, ESG_Level, Upturn) %>%
  summarize(Mean_5Year_Return = mean(`Close price`)) %>%
  ggplot(aes(x = Upturn, y = Mean_5Year_Return, fill = ESG_Level)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Period", y = "Closing Price", title = "Company Performance") +
  theme_bw() +
   scale_fill_manual(values = c("red","blue"))


```

H2a: Companies with higher ESG scores showcase superior financial performance during economic upturns compared to those with lower scores within the S&P 500, FTSE 250, and N225.

```{r}
aov(`Close price` ~ Stock * ESG_Level * Upturn , data = data_clean) %>% 
  summary()
```

# Modeling
## Regression analysis
```{r}
#Regression analysis
# Regression model not control for sector 
Model1 = lm(log(`Close price`)~`Refinitiv - ESG Score` + Country + Year, 
            data = data_clean)
# Regression model controlling for sector 
Model2 = lm(log(`Close price`)~`Refinitiv - ESG Score` + Country + Year + 
              `TRBC Business Sector`, data = data_clean)

modelsummary::modelsummary(list(Model1,Model2),
                            statistic = c("p.value"))
```


```{r}
# model validation
plot(Model2)
```

## DID

```{r}
# Load necessary packages
#install.packages("plm")
library(plm)

data_clean$ESG_Level = ifelse(data_clean$ESG_Level == 'High',1,0) 
data_clean$Upturn = ifelse(data_clean$Year =='Upturn',1,0)

# Run DiD regression
Model3 <- plm(log(Close.price) ~ TRBC.Business.Sector + ESG_Level + Upturn +
                ESG_Level:Upturn ,data = data_clean, 
              model = "pooling")


# Display results
modelsummary::modelsummary(Model3,
                            statistic = c("p.value"))

```



